---
title: "My Favorite / Most Used R and RStudio Stuff" 
subtitle: "I'd love to learn about yours!" 
author: "Rebecca Amodeo"
date: "`r format(Sys.time(), '%X, %d %B, %Y')`"
output:
  html_document:
    theme: journal
    toc: true
    toc_float:
      collapsed: false 
    number_sections: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
library(pacman) 
```
# base::  base R  
These are not exhaustive lists, just my most-used.   

## base operands  

 `==` means is equal to / is exactly in R;  
 
 `!=` means not equal to in R;  
 
 `>=` means greater than or equal to in R;  

## base functions  

## RStudio Shortcuts  
To comment out or in any line or set of lines.  

Step 1: highlight text / rows.  

Step 2: hold down `ctrl` + `shift` + hit `C`.  

Step 3: be so happy at all the time you saved :)  
```{r sample01} 
# sqlQuery(odbcConnect("StudentODS_in_Prod"),
  #              "SELECT A.RptChkDigitStdntID,
  #                     A.EthnicCd,
  #                     A.EconDsvntgFg,
  #                     A.LEPFg,
  #                     A.SpEdFg,
  #                     B.ChkDigitStdntID,
  #                     B.FinalChkDigitStdntID
  #                   FROM PIE.dbo.V_RO_SpringMbrshp A
  #                   LEFT JOIN [StudentODS].[dbo].[SSIDResolutionList] B
  #                     ON A.RptChkDigitStdntID = B.FinalChkDigitStdntID
  #                   WHERE Collection_ID = 2501 --3rd Period ADM in report year
  #                     AND A.EndDt IS NULL
  #                     AND B.EndDt IS NOT NULL") 
```

# here::

the here package saves time and code-line-space and allows both importing from and writing to folders in the open project's working directory.  

## here()
use the here() function from here::.  
You can see in the examples below the time, accuracy, and hand-cramp saving benefits of `here::here()` and imagine how much more is saved with this function when Working Directory file paths are long.  

**Example:**  

Working Directory (WD) == "C:/MyDocs/SampleProj"   

Folders in WD == c("data", "scripts", "output")   

**Import csv easily by nesting here() in read_csv():**  
`mydf <- read_csv(here("data", "fileanmemustincludethedotcsv.csv"))` This will create an R object in the global environment out of  "filenamemustincludethedotcsv.csv" from the folder "data" in the assigned/open working directory.  


Without `here::here()`, it looks like this: `mydf <- read_csv("C:/MyDocs/SampleProj/data/filenamemustincludethedotcsv.csv")`

```{r herehere1, eval = FALSE, echo = TRUE}
mydf <- read_csv(here("data", 
                      "fileanmemustincludethedotcsv.csv")) 

 # with both here::here() and magrittr's pipe operand 
 # to go right into data processing 
 ## note only one (1) occurrence of "mydf" needed for all this work! 
 ### insert googly-eyed emoji here 
mydf <- read_csv(here("data", 
                      "fileanmemustincludethedotcsv.csv")) %>%
  clean_names() %>%
  select(1:12)

 # without here::here() and magrittr's pipe.  
 ## note five (5) occurrences of "mydf" in the long-form. 
mydf <- read_csv("C:/MyDocs/SampleProj/data/filenamemustincludethedotcsv.csv")  
mydf <- clean_names(mydf)  
mydf <- select(mydf, 1:12)
```

**Write csv easily by nesting here() in write_csv():** 
`write_csv(RObjectIMade, here("output", "fileanmemustincludethedotcsv.csv"))` This writes my named R object from the global environment as "filenamemustincludethedotcsv.csv" to the folder "output" in the assigned/open working directory.   

Without here::here(), it looks like this: `write_csv(RObjectIMade, "C:/MyDocs/SampleProj/data/filenamemustincludethedotcsv.csv")`  

```{r herehere2, eval = FALSE, echo = TRUE}
write_csv(RObjectIMade, here("output",
                             "fileanmemustincludethedotcsv.csv"))  

 # with both here::here() and magrittr's pipe operand 
RObjectIMade %>% 
  write_csv(here("output",
                 "fileanmemustincludethedotcsv.csv"))

 # without here::here()  
write_csv(RObjectIMade,
          "C:/MyDocs/SampleProj/data/filenamemustincludethedotcsv.csv") 
```

# magrittr::  
I use this primarily for the pipe operand, `%>%`, which is now technically available in `base` R. I prefer the magrittr version and always load `magrittr` for the pipe.  

## %>% -- the pipe 
R package developers are so adorably nerdy. I love it. This package name is inspired by French painter Rene' Magritte. The flagship operand of this package is `%>%` named `the pipe` as an homage to Magritte's famous piece commenting on the nature of art and related abstractions, *The Treachery of Images*.  

# readr::  
I use this for many read and write functions that are more sophisticated than those available in `base` R.  

## read_csv()  

This function  

## read_xlsx()  

This function  

# RODBC::  
A life-saver for ODE analysts! 
## odbcCloseAll()  
This closes the connection opened to run the query. closing after each query run is best practice for this system.  
```{r RODBC1, eval = FALSE, echo = TRUE}
p_load(RODBC)

odbcCloseAll() # just add this after your sqlQuery() code. 
```

## sqlQuery()  
```{r RODBC2, eval = FALSE, echo = TRUE}
p_load(RODBC)

 # Sample SQL database query via RODBC.
EL <- sqlQuery(odbcConnect("ODBCConnectionName"),
               "SELECT * 
               FROM dbo.LEP 
               WHERE Collection_ID = 2498")
odbcCloseAll()
```

# skmir::  

my fav function(s) in this package 

## skim()  


# Other Quick Reference Guides  
I love the pretty "cheat sheets" for packages or thematically grouped operators and functions like these below.   

*You can get them all and more on the [RStudio cheatsheets page](https://www.rstudio.com/resources/cheatsheets/)*  

The `dplyr` [data transformation](https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-transformation.pdf) cheat sheet.    

The `ggplot2` [data viz](https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-visualization.pdf) cheat sheet.   

The `lubridate` [dates and times](https://raw.githubusercontent.com/rstudio/cheatsheets/main/lubridate.pdf) cheat sheet.    

The `plumber` [Web API](https://raw.githubusercontent.com/rstudio/cheatsheets/main/plumber.pdf) cheat sheet.    

The `purr` [apply functions](https://raw.githubusercontent.com/rstudio/cheatsheets/main/purrr.pdf) cheat sheet.  

The [R Markdown](https://raw.githubusercontent.com/rstudio/cheatsheets/main/rmarkdown.pdf) cheat sheet.  

The [RStudio IDE](https://raw.githubusercontent.com/rstudio/cheatsheets/main/rstudio-ide.pdf) cheat sheet.  

The `stringr` [string manipulation](https://raw.githubusercontent.com/rstudio/cheatsheets/main/strings.pdf) cheat sheet.  