---
title: "My Favorite / Most Used R and RStudio Stuff" 
subtitle: "I'd love to learn about yours!" 
author: "Rebecca Amodeo"
date: "`r format(Sys.time(), '%X, %d %B, %Y')`"
output:
  html_document:
    theme: journal
    toc: true
    toc_float:
      collapsed: false 
    number_sections: true 
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
library(pacman) 
```

# Doc Setup  
This is not an exhaustive list, just a small, curated presentation of my most-beloved `R` and `RStudio` operands, shortcuts, packages, functions, and more.  
  - A `package` is a set of functions and/or operands. Some packages have all "new" functions/operands (think *made from scratch*); others have a mix of "new" functions/operands and "legacy" (think *from another package*), and yet others are like awesome little coding museums that curate sets of legacy functions/operands that work well together for specific types of data processing (including designing and printing output, etc).   
  - A `function` is a   ...   
  - An `operand` is a   ...   

# base::  base R  


## base operands  
 `==` means is equal to / is exactly in `R`;  
 `!=` means not equal to in `R`;  
 `>=` means greater than or equal to in `R`;  

## base functions  

### duplicated()   
mutate(hasrepeat = duplicated((forever_prog_id)))  

## RStudio Shortcuts  
To comment out or in any line or set of lines.  
Step 1: highlight text / rows.  
Step 2: hold down `ctrl` + `shift` + hit `C`.  
Step 3: be so happy at all the time you saved :)  
```{r sample01, eval = FALSE, echo = TRUE} 
# p_load(RODBC) 
# 
# sqlQuery(odbcConnect("SQLConnection"),
  #              "SELECT A.RptChkDigitStdntID,
  #                     A.EthnicCd,
  #                     A.EconDsvntgFg,
  #                     A.LEPFg,
  #                     A.SpEdFg,
  #                     B.ChkDigitStdntID,
  #                     B.FinalChkDigitStdntID
  #                   FROM PIE.dbo.V_RO_SpringMbrshp A
  #                   LEFT JOIN [StudentODS].[dbo].[SSIDResolutionList] B
  #                     ON A.RptChkDigitStdntID = B.FinalChkDigitStdntID
  #                   WHERE Collection_ID = 2501 --3rd Period ADM in report year
  #                     AND A.EndDt IS NULL
  #                     AND B.EndDt IS NOT NULL") 
# 
# odbcCloseAll()
```

# dplyr::  
Where would I be without `dplyr`? In a sad place, that's where.  

## mutate()  
I use `mutate()` multiple times a day. :) Make a new column; overwrite an existing column. It's up to you!  
```{r dplyrmutate1, eval = FALSE, echo = TRUE}
toverify_1S1d <- teststudent %>%
  filter(PrimProgFg == "Y") %>%
  mutate(incohort = if_else(RptChkDigitStdntID %in% cohort4yr$ChkDigitStdntID, 
                            1, 0)) %>%
  mutate(goodassignment = as.factor(if_else(incohort == 1 & CTEEnrlmntSts == "Concentrator" & P5_1S1_D == "Y", "goodY",
                                  if_else(incohort == 1 & CTEEnrlmntSts == "Concentrator" & P5_1S1_D == "N", "ut-oh",
                                          if_else(incohort == 1 & CTEEnrlmntSts != "Concentrator" & P5_1S1_D == "N", "goodN",
                                          if_else(incohort == 0 & P5_1S1_D == "N", "goodN", "ut-oh"))))))
```


# here::  
The `here` package saves time and code-line-space and allows both importing from and writing to folders in the open working directory.  

## here()  
use the here() function from here::.  
You can see in the examples below the time, accuracy, and hand-cramp saving benefits of `here::here()` and imagine how much more is saved with this function when Working Directory file paths are long.  

**Example:**  
Working Directory (WD) == "C:/MyDocs/SampleProj"   
Folders in WD == c("data", "scripts", "output")   

**Import csv easily by nesting here() in read_csv():**  
`mydf <- read_csv(here("data", "fileanmemustincludethedotcsv.csv"))` This will create an R object in the global environment out of  "filenamemustincludethedotcsv.csv" from the folder "data" in the assigned/open working directory.  
Without `here::here()`, it looks like this: `mydf <- read_csv("C:/MyDocs/SampleProj/data/filenamemustincludethedotcsv.csv")`
```{r herehere1, eval = FALSE, echo = TRUE} 
p_load(here, janitor, readr, magrittr) 

mydf <- read_csv(here("data", 
                      "fileanmemustincludethedotcsv.csv")) 
```

With both `here::here()` and magrittr's `pipe operand`, we can  go right into data processing.  
Sweet, right? **note** only one (1) occurrence of `mydf` needed for all this work! Insert googly-eyed emoji here.  
```{r herehere2, eval = FALSE, echo = TRUE}
mydf <- read_csv(here("data", 
                      "fileanmemustincludethedotcsv.csv")) %>%
  clean_names() %>%
  select(1:12)
```

Without `here::here()` and magrittr's `pipe`. **note** five (5) occurrences of `mydf` in the long-form.  
```{r herehere3, eval = FALSE, echo = TRUE}
mydf <- read_csv("C:/MyDocs/SampleProj/data/filenamemustincludethedotcsv.csv")  
mydf <- clean_names(mydf)  
mydf <- select(mydf, 1:12)
```

**Write csv easily by nesting here() in write_csv():** 
`write_csv(RObjectIMade, here("output", "fileanmemustincludethedotcsv.csv"))` This writes my named R object from the global environment as "filenamemustincludethedotcsv.csv" to the folder "output" in the assigned/open working directory.   
Without here::here(), it looks like this: `write_csv(RObjectIMade, "C:/MyDocs/SampleProj/data/filenamemustincludethedotcsv.csv")`  
```{r herehere4, eval = FALSE, echo = TRUE}
write_csv(RObjectIMade, here("output",
                             "fileanmemustincludethedotcsv.csv"))  

 # with both here::here() and magrittr's pipe operand 
RObjectIMade %>% 
  write_csv(here("output",
                 "fileanmemustincludethedotcsv.csv"))

 # without here::here()  
write_csv(RObjectIMade,
          "C:/MyDocs/SampleProj/data/filenamemustincludethedotcsv.csv") 
```

# janitor::  
About this package...   

## clean_names()   
`clean_names()` gives you the best of both worlds with minimal friction in the transitions.  
I can use `SnakeCase` or `labels with spaces` in / to match client data systems but have R-processing-friendly 
```{r janitorcleannames1, eval = FALSE, echo = TRUE}

```

# magrittr::  
I use this primarily for the pipe operand, `%>%`, which is now technically available in `base` R. I prefer the magrittr version and always load `magrittr` for the pipe.  

## %>% -- the pipe operand  
R package developers are so adorably nerdy. I love it. This package name is inspired by French painter Rene' Magritte. The flagship operand of this package is `%>%` named `the pipe` as an homage to Magritte's famous piece commenting on the nature of art and related abstractions, *The Treachery of Images*.  

# pacman::  
An early and enduring fav! This package applies function generally limited to one (1) `x` argument to multiple specified objects.  

## p_load()  

My number one in this package is `p_load`. I use it in virtually all my scripts.  

This performs **both** `base::install.package()` **and** `base::library()` **as-freakin-needed**. Massive time and sanity saver for me.  
```{r pacmanpload}

```


# readr::  
I use this for many read and write functions that are more sophisticated than those available in `base` R.  

## read_csv()  

This function  

## read_xlsx()  

This function  

# RODBC::  
A life-saver for ODE analysts!  
The cran entry is [here](https://cran.r-project.org/web/packages/RODBC/index.html).  

I answered a stackoverflow question using this package [here](https://stackoverflow.com/questions/60536206/speed-up-odbcdbfetch/73880848#73880848).  

## odbcCloseAll()  
This closes the connection opened to run the query. closing after each query run is best practice for this system.  
```{r RODBC1, eval = FALSE, echo = TRUE}
p_load(RODBC)

odbcCloseAll() # add this after your sqlQuery() code. 
```

## sqlQuery()  
```{r RODBC2, eval = FALSE, echo = TRUE}
p_load(RODBC)

 # Sample SQL database query via RODBC.
EL <- sqlQuery(odbcConnect("ODBCConnectionName"),
               "SELECT * 
               FROM dbo.table 
               WHERE Collection_ID = 2498")
odbcCloseAll()
```

# skmir::  
My fav function(s) in this package  

## skim()  

# YAML stuff for .Rmd  
I have some go-to YAML formatting of my .Rmd output.   

> title: "My Favorite / Most Used R and RStudio Stuff"    
> subtitle: "I'd love to learn about yours!"   
> author: "Rebecca Amodeo"  
> date: "`r format(Sys.time(), '%X, %d %B, %Y')`"  
> output:  
  >>  html_document:  
  >>  theme: journal  
  >>  toc: true  
  >>  toc_float:  
  >>  (indent)collapsed: false   
  >>  number_sections: true   

match.arg(theme, themes()) : 
  'arg' should be one of "default", "bootstrap", "cerulean", "cosmo", "darkly", "flatly", "journal", "lumen", "paper", "readable", "sandstone", "simplex", "spacelab", "united", "yeti"  
  
I often theme by project to help myself and teammates keep track of which we are working on any given moment. Examples:  

  - This document uses `journal`   
  - Secondary Career Pathways uses `lumen`   
  - ADI_90pct_Validation uses `darkly`  
  - Civil Rights Report uses `sandstone`  


# zoomGroupStats()   
I haven't gotten this yet. Documentation at http://zoomgroupstats.org/  

```{r} 
p_load(zoomGroupStats)

zchat <- processZoomChat(
  fname = system.file(here("data", "meeting_saved_chat.txt"),
                      package = "zoomGroupStats"),
  sessionStartDateTime = "2022-10-14 11:55:17",
  languageCode = "en") 

```


# Other Quick Reference Guides  
I love the pretty "cheat sheets" for packages or thematically grouped operators and functions like these below.   
*You can get them all and more on the [RStudio cheatsheets page](https://www.rstudio.com/resources/cheatsheets/)*  
The `dplyr` [data transformation](https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-transformation.pdf) cheat sheet.    
The `ggplot2` [data viz](https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-visualization.pdf) cheat sheet.   
The `lubridate` [dates and times](https://raw.githubusercontent.com/rstudio/cheatsheets/main/lubridate.pdf) cheat sheet.    
The `plumber` [Web API](https://raw.githubusercontent.com/rstudio/cheatsheets/main/plumber.pdf) cheat sheet.    
The `purr` [apply functions](https://raw.githubusercontent.com/rstudio/cheatsheets/main/purrr.pdf) cheat sheet.  
The [R Markdown](https://raw.githubusercontent.com/rstudio/cheatsheets/main/rmarkdown.pdf) cheat sheet.  
The [RStudio IDE](https://raw.githubusercontent.com/rstudio/cheatsheets/main/rstudio-ide.pdf) cheat sheet.  
The `stringr` [string manipulation](https://raw.githubusercontent.com/rstudio/cheatsheets/main/strings.pdf) cheat sheet.  